---
title: "Dotted charts"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      print: false
---

```{r include = F}
library(tidyverse)
library(bupaR)
library(pander)
library(lubridate)
```
```{r echo = F}
htmltools::includeHTML("tracking_google_analytics.html")
```

Dotted charts can be made with the `dotted_chart` function. A dotted chart is a graph in which each activity instance is displayed with a point. The x-axis referce to the time aspect, while the y-axis refers to cases. The dotted chart function has 3 arguments

* x: Either absolute (absolute time on x-axis) or relative (time difference since start case on x-axis)
* y: The ordering of the cases along the y-axis: by start, end, or duration.
* color: The attribute used to color the activity instances. Default to the activity type.

## Static dotted charts

Below, you can see some examples for dotted charts with various configurations

```{r}
patients %>%
	dotted_chart(x = "absolute", y = "start")
patients %>%
	dotted_chart(x = "absolute", y = "start", color = "employee")
patients %>%
	dotted_chart(x = "relative", y = "duration", color = "employee")
```

## Interactive dotted charts

The function `idotted_chart` opens a shiny app which can be used to modify the arguments of the dotted chart interactively.
