---
title: eventdataR
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      print: false
---
	
```{r echo = F}
htmltools::includeHTML("tracking_google_analytics.html")
library(bupaR)
```
The `eventdataR` package contains several real-life and artificial event logs. Each can be loaded using the `data` function. The currently available event logs are listed below. More event logs will be added in the future.

## Artifical data

### Patients



Artificial eventlog about patients arriving in an emergency department of a hospital. This event log was used as the running example in the journal paper entitled [Retrieving batch organisation of work insights from event logs](http://www.sciencedirect.com/science/article/pii/S0167923617300374).


```{r message = F, warning = F}
library(eventdataR)
patients %>% summary
```

## Real-life data

### Sepsis

```{r}
sepsis
```


### Hospital Log

```{r}
hospital
```


### Hospital Billing

```{r}
hospital_billing
```



### Road Traffic Fine Management

```{r}
traffic_fines
```



	
	
	