---
title: Process Matrix
---


```{r}
library(bupaR)
```


## Introduction
A precendence diagrams is a two-dimensional matrix showing the flows between activities. It can be contain different type values, by adjusting the type argument.

*	Absolute frequency of flows
*	Relative frequency of flows
*	Relative frequency of flows, for each antecendent
	*	I.e. given antecendent A, it is followed x% of the time by Consequent B
*	Relative frequency of flows, for each consequent
	*	I.e. given consequent B, it is preceded x% of the time by Antecedent A
	
The precedence diagrams can be visualized using the generic plot function. Below, an example of each of the different types is shown.

##Examples

### Absolute Frequencies

```{r}

patients %>%
	precedence_matrix(type = "absolute") 
patients %>%
	precedence_matrix(type = "absolute") %>%
	plot
```


### Relative Frequencies
```{r}
patients %>%
	precedence_matrix(type = "relative") %>%
	plot
```


### Antecedent-wise Frequencies
```{r}
patients %>%
	precedence_matrix(type = "relative-antecedent") %>%
	plot
```


### Consequent-wise Frequencies
```{r}
patients %>%
	precedence_matrix(type = "relative-consequent") %>%
	plot
```	


	
	
	
	